Seeing Double: Transient Dips and Photometric Binarity in PTFO$\,$8-8695

### 2020-04-21 paper/ms.tex
added words

One particularly interesting feature of the dip is its epoch.  Over
the past decade, many investigators have observed \ptfo.  Its dips do
not always occur on a perfectly linear ephemeris
\citep{yu_tests_2015}.  In fact, \citet{tanimoto_evidence_2020}
recently provided stark evidence for different behavior altogether:
over a time-span of years, the dip ``splits'' into distinct groups at
particular repeating phases.  See for instance their Figures~2
through~4.  Fitting a decade of observations, they provided the
following linear ephemeris.
\begin{align}
t_0\ {\rm BJD}_{\rm TDB} &= 2455543.943 \pm 0.002 \\
P &= 0.4483993 \pm 0.0000006 d.
\end{align}

@@

%FIXME write it from here!

If there were a giant planet transiting \ptfo, it would tidally
distort the host star, and cause ellipsoidal photometric modulations.
The amplitude of the ellipsoidal distortion for a $1\,M_{\rm Jup}$
companion would be about 1400$\,$ppm
\citep{shporer_astrophysics_2017}.  This is significantly larger than
the typical ellipsopidal modulation induced by close-in giant planets
because the host star is puffy, and still on the pre-main-sequence.
For our estimate, we assumed $R_\star = 1.39 R_\odot$, and $M_\star =
0.39 M_\odot$ \citep{van_eyken_ptf_2012}.

Our preferred model does detect a significant ellipsoidal signal,
parametrized as the ``$B_1$'' component.  The amplitude of the signal
is $0.53 \pm 0.06\%$ (see Table~2).  Interpreted as being caused by a
planet, it would imply a minimum planet mass $M_{\rm p} \sin i$ of
$3.8\,M_{\rm Jup}$.

%FIXME but argue against

\subsection{Similar Lightcurves}

Figure~\ref{fig:brethren}

@@

%FIXME: the PTFO lightcurve looks so much worse! Why is it less dense??
\begin{figure*}[hbtp]
	\begin{center}
		\leavevmode
		\includegraphics[width=1\textwidth]{f7.pdf}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf \ptfo\ and its brethren.}
    Five transient and persistent flux dip stars, selected based on
    their visual similarity to the short-period signal in \ptfo, are
    as follows.
  	EPIC 204143627, 1.1250d, dips change depth too
  	EPIC 204321142  Usco flux dip, 0.476d
  	EPIC 204270520  USco flux dip, 0.512d
  	% EPIC 205046529B, 1.8358d, indicates dips can be at weird phases too
  	RIK-210 = EPIC 205483258
 	EPIC 204787516  USco flux dip or possible? EB, 0.487d
 	We found these objects through studies by \citet{stauffer_orbiting_2017}, \citet{david_transient_2017}, and \citet{rebull_usco_2018}.
		\label{fig:brethren}
	}
\end{figure*}

>>>>

One particularly interesting feature of the dip is its epoch.  Over
the past decade, many investigators have observed \ptfo.  Its dips do
not always occur on a perfectly linear ephemeris
\citep{yu_tests_2015}.  In fact, \citet{tanimoto_evidence_2020}
recently provided stark evidence for different behavior altogether:
over a time-span of years, the dip ``splits'' into distinct groups at
particular repeating phases.  See for instance their Figures~2
through~4.  Fitting a decade of observations, they provided the
following linear ephemeris, which we did not find any need to update.
\begin{align}
t_0\ {\rm BJD}_{\rm TDB} &= 2455543.943 \pm 0.002 \\
P &= 0.4483993 \pm 0.0000006 d.
\end{align}

@@

Visually, the out-of-dip modulation at the 10.74$\,$hour period
resembles a slightly asymmetric sinusoid (Figure~\ref{fig:phasefold}).
Non-zero contributions in both the first and second harmonics are
detected (Table~2).  The first sine and cosine harmonic both have
amplitudes of roughly $0.9\pm0.1\%$.  The second sine harmonic has
amplitude $0.16 \pm 0.07\%$, so is non-zero at a significance of only
2.3$\sigma$.  The second cosine harmonic has negative amplitude $0.53
\pm 0.06\%$.  In our sign convention, the fact that it is negative
means that this component peaks at phase 0.25 and 0.75.

\subsubsection{Planetary interpretation}
If there were a giant planet transiting \ptfo, it would tidally
distort the host star, and cause ellipsoidal photometric modulations
that also peak at quadrature \citep[see][]{shporer_astrophysics_2017}.
Interpreting the second cosine harmonic as being caused by a planet,
it would imply a minimum planet mass $M_{\rm p} \sin i$ of
$3.8\,M_{\rm Jup}$.  For this estimate, we assumed $R_\star = 1.39
R_\odot$, and $M_\star = 0.39 M_\odot$ \citep{van_eyken_ptf_2012}.
This ellipsoidal amplitude is larger than the typical
modulations induced by close-in giant planets because the host star is
puffy, and still on the pre-main-sequence.

The planetary interpretation however does not readily explain the
first sine and cosine harmonics.  Interpreting the sine component as
Doppler beaming would imply a secondary mass greater than the primary
($0.86\,M_\odot$).  Interpreting the cosine component as reflected or
emitted light from the planet's surface is nonsensical because the sign is wrong---the
planet would need to be {\it absorbing} light.

\subsubsection{Similar Lightcurves}

\begin{figure*}[hbtp]
	\begin{center}
		\leavevmode
		\includegraphics[width=1\textwidth]{f7.pdf}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf \ptfo\ and its brethren.}
    Five transient and persistent flux dip stars selected based on
    their visual similarity to the short-period signal in \ptfo are as
    shown.  They include EPIC 204143627, EPIC 204321142, EPIC
    204270520, EPIC 205483258 (RIK-210), and EPIC 204787516.  RIK-210
    has the longest period of any of these objects.  EPIC 204787516
    has a less angular dip, and could also be an eclipsing binary.
    All stars have similar properties as \ptfo, but are in Upper Sco.
    We found these objects through studies by
    \citet{stauffer_orbiting_2017}, \citet{david_transient_2017}, and
    \citet{rebull_usco_2018}.
		\label{fig:brethren}
	}
\end{figure*}

When physical explanations are not immediately clear, taxonomy can be
a useful exercise.  To our knowledge, of order 10 lightcurves with
similar morphologies have been seen in surveys of low-mass weak-lined
T Tauri stars in regions including $\rho$ Oph, Upper Sco, Taurus, and perhaps
the Pleiades
\citep{rebull_rotation_2016,david_transient_2017,stauffer_orbiting_2017,stauffer_rotevol_2018,rebull_usco_2018,rebull_rotation_2020}.
These surveys were performed using K2 \citep{howell_k2_2014}.  We
downloaded a few of these lightcurves from MAST, opting for the
EVEREST reductions \citep{luger_everest_2016,luger_update_2018}.  They
are plotted in Figure~\ref{fig:brethren}

These lightcurves have been classified phenomenologically as
``persistent flux dips'' or ``transient flux dips'', based on whether
their depths and durations show variability over the 90-day K2
campaigns \citep{stauffer_orbiting_2017}.
This class of objects has the following properties.
\begin{enumerate}
  \item Weak-lined T Tauri stars.
  \item Spectral type typically M2-M5 ({\it e.g.},
    \citet{rebull_usco_2018}~Figure~20).
  \item Ages $\lesssim$ 100$\,$Myr.
  \item Stars typically lack spectroscopic accretion indicators 
  \item Phased lightcurves show shallow, angular dips, usually
    superposed on large-amplitude smooth variability (interpreted as
    stellar rotation).
  \item Rotation is usually rapid ($P<1\,{\rm day}$).
  \item Rarely any infrared excess (never any W4 detection; only a few
    cases with W3 detctions).
  \item Can show multiple dips per cycle ({\it e.g.}, EPIC 203692610,
    RIK-210).
  \item Dip depths, durations, and phasescan vary over timescales of days
    ({\it e.g.}, EPIC 204143627).
  \item Dip depths can change after flares ({\it e.g.}, RIK-210.
  \item Rare ($\lesssim 1\%$ of relevant stars) at a population level
    \citep{rebull_usco_2018}.
\end{enumerate}
The 10.74 hour signal in \ptfo\ meets essentially all of these
criteria.

@@
